---
title: "JSC370 Final Project - Identify and Monitor the Most Phosphorous-Polluted Locations in the Great Lakes"
author: "Steven Liu"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

## Description of the Project

This project focuses on investigating the water quality of the Great Lakes region in North America, with a specific focus on phosphorus levels in the water. This project mainly uses two datasets: 1) The water quality dataset from [The Canadian Open Data Portal](https://data-donnees.ec.gc.ca/data/substances/monitor/great-lakes-water-quality-monitoring-and-aquatic-ecosystem-health-data/great-lakes-water-quality-monitoring-and-surveillance-data/?lang=en), and 2) The city dataset, obtained through web scraping from [Latlong](https://www.latlong.net/category/cities-236-15.html). They explored and visualized the data using time series plots, boxplots, and interactive maps. The results of their analysis showed that Lake Erie and Lake Ontario have the highest phosphorus levels and are facing more severe pollution issues than the other Great Lakes. They also identified specific cities with the worst water quality, such as Hamilton, ON, Canada; Niagara Falls, NY, USA; Toledo, OH, USA; Sandusky, OH, USA; and Cleveland, OH, USA. Furthermore, they have investigated whether there is evidence of improvement in water quality at the identified locations from 2000 to the present. they have found no evidence of improvement and, in fact, have identified a worsening trend in water quality over time, especially for the west-coast cities. Overall, this research paper provides valuable insights into the water quality issues in the Great Lakes region and can inform future efforts to improve water quality in the area.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
library(car)
library(zoo)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(leaflet)
library(rvest)
library(xml2)
library(patchwork)
library(leaflet.esri)
library(shiny)
library(htmlwidgets)
library(data.table)
library(plotly)
library(knitr)
library(widgetframe)
```

```{r}
# Read Data
water1 <- read.csv("data/Lake Ontario_Water Quality_2000-present.csv")
water2 <- read.csv("data/Lake Erie_Water Quality_2000-present.csv")
water3 <- read.csv("data/Lake Superior_Water Quality_2000-present.csv")
water4 <- read.csv("data/Lake Huron_Water Quality_2000-present.csv")
water5 <- read.csv("data/Georgian Bay_Water Quality_2000-present.csv")
# Bind data for great lakes together
water <- rbind(water1, water2, water3, water4, water5)
```

```{r, eval=FALSE}
# Add Canada Cities
city <- html_table(xml2::xml_find_all(
  read_html(x ="https://www.latlong.net/category/cities-40-15.html"),
          xpath = "/html/body/main/table"))[[1]]
for (i in 2:4){
  html <- paste0(
    substring("https://www.latlong.net/category/cities-40-15-2.html", 1, 45),
    '-', i,'.html')
  city <- bind_rows(city, html_table(xml2::xml_find_all(
    read_html(x = html),
    xpath = "/html/body/main/table"))[[1]])
}
```

```{r, eval=FALSE}
# Add U.S. Cities
city <- bind_rows(city, html_table(xml2::xml_find_all(
  read_html(x ="https://www.latlong.net/category/cities-236-15.html"),
          xpath = "/html/body/main/table"))[[1]])

for (i in 2:12){
  html <- paste0(
    substring("https://www.latlong.net/category/cities-236-15-2.html", 1, 46),
    '-', i,'.html')
  city <- bind_rows(city, html_table(xml2::xml_find_all(
    read_html(x = html),
    xpath = "/html/body/main/table"))[[1]])
}
```

```{r, eval=FALSE}
# Add additional cities
additional_city <- read.csv("data/Additional_cities.csv")
additional_city <- additional_city %>%
  mutate(`Place Name` = Place.Name) %>% 
  select(-Place.Name)

city <- bind_rows(city, additional_city)
```

```{r}
water_reduced <- water %>% 
  # Rename main variables, and create separate column for year and month
  mutate(lake = ifelse(WATER_BODY == "GEORGIAN BAY", "LAKE HURON", WATER_BODY),
         station = paste0(WATER_BODY," ",as.character(as.integer(PSN))), 
         year = as.integer(substr(STN_DATE, start = 7, stop = 10)),
         month = as.integer(substr(STN_DATE, start = 4, stop = 5)),
         yearmon = as.yearmon(paste0(year, "-", month)),
         range = case_when(year <= 2005 ~ "2000-2005",
                           (year > 2005 & year <= 2010) ~ "2005-2010",
                           (year > 2010 & year <= 2015) ~ "2010-2015",
                           (year > 2015 & year <= 2020) ~ "2015-2020",
                           TRUE ~ "2020-present"),
         hazard_name = FULL_NAME,
         unit = UNITS,
         value = VALUE,
         lat = LATITUDE_DD,
         # We notice that the lon is actually - LONGITUDE
         lon = - LONGITUDE_DD) %>% 
 # Select main variables of interest
  select(lake, station, year, month, yearmon, range, hazard_name, value, lat, lon) 
```

```{r}
# Remove NA
water_reduced <- na.omit(water_reduced)
water_reduced <- water_reduced %>% filter(value > 0)

# As we notice that the lat and lon for each station is different, we want to use mean as their lat and lon
water_reduced <- water_reduced %>% group_by(station) %>% mutate(lat = round(mean(lat), 3),
                                               lon = round(mean(lon), 3))
# Remove duplicate
water_reduced <- unique(water_reduced)
```


```{r}
# Select the phosphorous 
water_phos <- water_reduced %>% 
  filter(hazard_name == "PHOSPHOROUS,TOTAL" | hazard_name == "TOTAL PHOSPHOROUS") %>% 
  mutate(level = case_when(value < 0.1 ~ "Good", 
                           value >= 0.1 & value < 0.25 ~ "Poor",
                           TRUE ~ "Very Bad"))
# Check for missing/error values
summary(water_phos)
```

```{r}
# Add Canada Cities
city <- html_table(xml2::xml_find_all(
  read_html(x ="https://www.latlong.net/category/cities-40-15.html"),
          xpath = "/html/body/main/table"))[[1]]
for (i in 2:4){
  html <- paste0(
    substring("https://www.latlong.net/category/cities-40-15-2.html", 1, 45),
    '-', i,'.html')
  city <- bind_rows(city, html_table(xml2::xml_find_all(
    read_html(x = html),
    xpath = "/html/body/main/table"))[[1]])
}
```

```{r}
# Add U.S. Cities
city <- bind_rows(city, html_table(xml2::xml_find_all(
  read_html(x ="https://www.latlong.net/category/cities-236-15.html"),
          xpath = "/html/body/main/table"))[[1]])

for (i in 2:12){
  html <- paste0(
    substring("https://www.latlong.net/category/cities-236-15-2.html", 1, 46),
    '-', i,'.html')
  city <- bind_rows(city, html_table(xml2::xml_find_all(
    read_html(x = html),
    xpath = "/html/body/main/table"))[[1]])
}
```

```{r}
# Add additional cities
additional_city <- read.csv("data/Additional_cities.csv")
additional_city <- additional_city %>%
  mutate(`Place Name` = Place.Name) %>% 
  select(-Place.Name)

city <- bind_rows(city, additional_city)
```

```{r}
# Lake Cities
lake_city <- city %>% 
  filter(Latitude > city[`Place Name` == "Mansfield, OH, USA",]$Latitude)
lake_city <- lake_city %>% 
  filter(Latitude < lake_city[`Place Name` == "Fermont, QC, Canada",]$Latitude)
lake_city <- lake_city %>% 
  filter(Longitude > lake_city[`Place Name` == "Winkler, MB, Canada",]$Longitude)
lake_city <- lake_city %>% 
  filter(Longitude < lake_city[`Place Name` == "Ottawa, ON, Canada",]$Longitude)

lake_city <- lake_city %>% filter(`Place Name` != "Monroe, OH, USA")
```

```{r}
# For each station, find its nearest city
water_phos_with_city <- water_phos %>% 
  crossing(lake_city) %>% 
  mutate(distance = round(sqrt((lat - Latitude) ** 2 + (lon - Longitude) ** 2), 3))%>% 
  group_by(station) %>% 
  filter(distance == min(distance)) %>%
  mutate(city = `Place Name`,
         lon_city = Longitude,
         lat_city = Latitude) %>% 
  select(-`Place Name`, -Longitude, -Latitude)
```

```{r}
# Average Phosphorous per Lake
average_per_lake <- water_phos %>% group_by(lake, yearmon) %>% 
  summarise(mean_value = round(mean(value), 5), n = n())
```

```{r}
# Average Phosphorous per Lake adjusted
average_per_lake_adj <- water_phos %>% group_by(lake, year) %>% 
  summarise(mean_value = round(mean(value), 5), n = n())
```

```{r}
# Average Phosphorous per Lake Station Regardless on Time
average_per_lake_station <- water_phos_with_city %>% 
  group_by(lake, station, lon, lat, lon_city, lat_city, city) %>% 
  summarise(mean_value = round(mean(value), 5), 
            time_difference = max(yearmon) - min(yearmon)) %>%
  # We want to only consider station with data more than 3 years
  filter(time_difference > 3)
```

```{r}
# Average Phosphorous per Lake Station by Time Period
average_per_lake_station_with_period <- water_phos_with_city %>% 
  group_by(lake, station, lon, lat, city, lon_city, lat_city, range) %>% 
  summarise(mean_value = round(mean(value), 5), 
            n = n()) %>% 
  filter(station %in% average_per_lake_station$station)

```

```{r}
# Average Phosphorous per Lake - City
average_per_city <- water_phos_with_city %>% 
  group_by(lake, lon_city, lat_city, city) %>% 
  summarise(lon = round(mean(lon), 3),
            lat = round(mean(lat), 3),
            mean_value = round(mean(value), 5), 
            time_difference = max(yearmon) - min(yearmon)) %>%
  # We want to only consider station with data more than 3 years
  filter(time_difference > 3) %>% 
  filter(lake == "LAKE ONTARIO" | lake == "LAKE ERIE")

# 5 Most Polluted Cities per Lake 
top_5_city_per_lake <- average_per_city %>% 
  group_by(lake) %>% 
  slice_max(order_by = mean_value, n = 5)
```
```{r}
# Average Phosphorous per Lake Station by Time Period
average_per_city_with_period <- water_phos_with_city %>% 
  group_by(lake, lon_city, lat_city, city, range) %>% 
  summarise(lon = round(mean(lon), 3),
            lat = round(mean(lat), 3),
            mean_value = round(mean(value), 5), 
            n = n()) %>%
  filter(city %in% average_per_city$city) %>% 
  filter(lake == "LAKE ERIE" | lake == "LAKE ONTARIO")

# 5 Most Polluted Cities per Lake 
top_5_city_per_lake_with_period <- average_per_city_with_period %>% 
  group_by(lake, range) %>% 
  slice_max(order_by = mean_value, n = 5)

```

```{r}
plt_series <- ggplot(
  average_per_lake_adj, aes(x = year, y = mean_value, colour = lake)) + 
  geom_line() +
  theme_bw() + 
  ggtitle("Adjust Time Series for Mean Phosphorous Level of Great Lakes")
plotly_series <- ggplotly(plt_series)

```
```{r}
city_mat_table <- average_per_city_with_period[,c("city", "range", "mean_value")]  %>% 
  pivot_wider(names_from = city, values_from = mean_value) %>% 
  column_to_rownames("range") %>% select(-`St. Thomas, ON, Canada`)
city_mat <- as.matrix(city_mat_table)

plotly_heatmap <- city_mat_table %>% 
  plot_ly(
    # date
    x = rownames(city_mat_table),
    # states
    y = colnames(city_mat_table),
    # new_cases
    z = t(city_mat),
    type = "heatmap"
  )
```

```{r}
# 10 Most Polluted stations per Lake 
top_10_station_per_lake <- average_per_lake_station %>% 
  group_by(lake) %>% 
  slice_max(order_by = mean_value, n = 10)

top_10_station_per_lake
```

```{r}
# 10 Most Polluted stations per Lake 
top_10_station_per_lake_with_period <- average_per_lake_station_with_period %>% 
  group_by(lake, range) %>% 
  slice_max(order_by = mean_value, n = 10)

top_10_station_per_lake_with_period
```


```{r, include = TRUE}

color_scheme <- colorNumeric(c('green', 'red', 'darkred'),
                       domain = average_per_lake_station_with_period$mean_value)
  

interactive_leaflet1 <- leaflet() %>% 
  addProviderTiles('OpenStreetMap') %>% 
  # Default
  addPolylines(
    lng = c(-180, 180),
    lat = c(45, 45),
    dashArray = "10, 10",
    color = "red"
  ) %>% 
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station$station,
                            ": \n",
                            average_per_lake_station$mean_value),
             data = average_per_lake_station, 
             group = "2000-present") %>% 
  addAwesomeMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake, 
             popup = paste0(top_10_station_per_lake$station,
                            ": \n",
                            top_10_station_per_lake$mean_value), 
             group = "2000-present") %>% 
  # 2000-2005
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2000-2005",
             ]$station,
                            ": \n",
                            average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2000-2005",
             ]$mean_value),
             data = average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2000-2005",
             ], 
             group = "2000-2005") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2000-2005",
             ], 
             popup = paste0(top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2000-2005",
             ]$station,
                            ": \n",
                            top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2000-2005",
             ]$mean_value), 
             group = "2000-2005") %>% 
  # 2005-2010
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2005-2010",
             ]$station,
                            ": \n",
                            average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2005-2010",
             ]$mean_value),
             data = average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2005-2010",
             ], 
             group = "2005-2010") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2005-2010",
             ], 
             popup = paste0(top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2005-2010",
             ]$station,
                            ": \n",
                            top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2005-2010",
             ]$mean_value), 
             group = "2005-2010") %>% 
  # 2010-2015
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2010-2015",
             ]$station,
                            ": \n",
                            average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2010-2015",
             ]$mean_value),
             data = average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2010-2015",
             ], 
             group = "2010-2015") %>% 
   addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2010-2015",
             ], 
             popup = paste0(top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2010-2015",
             ]$station,
                            ": \n",
                            top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2010-2015",
             ]$mean_value), 
             group = "2010-2015") %>% 
  # 2015-2020
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2015-2020",
             ]$station,
                            ": \n",
                            average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2015-2020",
             ]$mean_value),
             data = average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2015-2020",
             ], 
             group = "2015-2020") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2015-2020",
             ], 
             popup = paste0(top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2015-2020",
             ]$station,
                            ": \n",
                            top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2015-2020",
             ]$mean_value), 
             group = "2015-2020")%>% 
  # 2020-present
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 1000, 
             popup = paste0(average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2020-present",
             ]$station,
                            ": \n",
                            average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2020-present",
             ]$mean_value),
             data = average_per_lake_station_with_period[
               average_per_lake_station_with_period$range == "2020-present",
             ], 
             group = "2020-present") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2020-present",
             ], 
             popup = paste0(top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2020-present",
             ]$station,
                            ": \n",
                            top_10_station_per_lake_with_period[
               top_10_station_per_lake_with_period$range == "2020-present",
             ]$mean_value), 
             group = "2020-present")%>% 

  addLegend("bottomleft", pal = color_scheme, 
            average_per_lake_station$mean_value,
            title = "Phosphorous (mg/L)", 
            opacity = 1) %>% 
  setView(lng = -81, lat = 45, zoom = 5.3) %>%
  # Layers control
  addLayersControl(
    baseGroups = c("2000-present", "2000-2005", "2005-2010",
                   "2010-2015", "2015-2020","2020-present"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


```{r, include = TRUE}
color_scheme <- colorNumeric(c('green', 'red', 'darkred'),
                       domain = average_per_lake_station_with_period$mean_value)

interactive_leaflet2 <- leaflet() %>% 
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city$city,
                            ": \n",
                            average_per_city$mean_value),
             data = average_per_city, group = "2000-present") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake, 
             popup = paste0(top_5_city_per_lake$city,
                            ": \n",
                            top_5_city_per_lake$mean_value), group = "2000-present") %>% 

  # 2000-2005
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city_with_period[
              average_per_city_with_period$range == "2000-2005",
             ]$city,
                            ": \n",
                           average_per_city_with_period[
              average_per_city_with_period$range == "2000-2005",
             ]$mean_value),
             data = average_per_city_with_period[
              average_per_city_with_period$range == "2000-2005",
             ], group = "2000-2005") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2000-2005",
              ], 
             popup = paste0(top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2000-2005",
              ]$city,
                            ": \n",
                            top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2000-2005",
              ]$mean_value), group = "2000-2005") %>%  
  
  # 2005-2010
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city_with_period[
              average_per_city_with_period$range == "2005-2010",
             ]$city,
                            ": \n",
                           average_per_city_with_period[
              average_per_city_with_period$range == "2005-2010",
             ]$mean_value),
             data = average_per_city_with_period[
              average_per_city_with_period$range == "2005-2010",
             ], group = "2005-2010") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2005-2010",
              ], 
             popup = paste0(top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2005-2010",
              ]$city,
                            ": \n",
                            top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2005-2010",
              ]$mean_value), group = "2005-2010") %>%  
  # 2010-2015
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city_with_period[
              average_per_city_with_period$range == "2010-2015",
             ]$city,
                            ": \n",
                           average_per_city_with_period[
              average_per_city_with_period$range == "2010-2015",
             ]$mean_value),
             data = average_per_city_with_period[
              average_per_city_with_period$range == "2010-2015",
             ], group = "2010-2015") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2010-2015",
              ], 
             popup = paste0(top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2010-2015",
              ]$city,
                            ": \n",
                            top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2010-2015",
              ]$mean_value), group = "2010-2015") %>%  
  
  # 2015-2020
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city_with_period[
              average_per_city_with_period$range == "2015-2020",
             ]$city,
                            ": \n",
                           average_per_city_with_period[
              average_per_city_with_period$range == "2015-2020",
             ]$mean_value),
             data = average_per_city_with_period[
              average_per_city_with_period$range == "2015-2020",
             ], group = "2015-2020") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2015-2020",
              ], 
             popup = paste0(top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2015-2020",
              ]$city,
                            ": \n",
                            top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2015-2020",
              ]$mean_value), group = "2015-2020") %>%  
  # 2020-present
  addCircles(lng = ~lon,
             lat = ~lat, 
             color = ~color_scheme(mean_value), 
             fillOpacity = 1, 
             radius = 3000, 
             popup = paste0(average_per_city_with_period[
              average_per_city_with_period$range == "2020-present",
             ]$city,
                            ": \n",
                           average_per_city_with_period[
              average_per_city_with_period$range == "2020-present",
             ]$mean_value),
             data = average_per_city_with_period[
              average_per_city_with_period$range == "2020-present",
             ], group = "2020-present") %>% 
  addMarkers(lat=~lat, 
             lng=~lon, 
             data = top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2020-present",
              ], 
             popup = paste0(top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2020-present",
              ]$city,
                            ": \n",
                            top_5_city_per_lake_with_period[
              top_5_city_per_lake_with_period$range == "2020-present",
              ]$mean_value), group = "2020-present") %>%  
  
  addLegend("bottomleft", pal = color_scheme, 
            average_per_lake_station$mean_value,
            title = "Phosphorous (mg/L)", 
            opacity = 1) %>% 
  setView(lng = -80, lat = 43, zoom = 6.2) %>%
  # Layers control
  addLayersControl(
    baseGroups = c("2000-present", "2000-2005", "2005-2010", "2010-2015",
                   "2015-2020", "2020-present"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


## Showcasing plots {.tabset}

The following are the four main interactive figures that display the results of this research:

Figure 1 illustrates the Adjusted Time Series for the Mean Phosphorous Levels of the Great Lakes, depicting the trend of phosphorous levels in the lakes from 2000 to the present.

Figure 2 (Leflet 1) presents the average phosphorous levels for each observation station (shown by dots) in all Great Lakes, highlighting the top 10 polluted stations using markers.

Figure 3 (Leflet 2) displays the average phosphorous levels for each city (represented by dots) near Lake Erie and Lake Ontario, labeling the top 5 polluted cities with markers.

Figure 4 shows the Heatmap for Phosphorous Levels over time by City, demonstrating the phosphorous levels of cities near Lake Erie and Lake Ontario from 2000 to the present.


### Figure 1


```{r, include=TRUE}
plotly_series
```



### Figure 2 (Leflet 1)


```{r, include=TRUE}
interactive_leaflet1
```
### Figure 3 (Leflet 2)


```{r, include=TRUE}
interactive_leaflet2
```

### Figure 4


```{r, include=TRUE}
plotly_heatmap <- plotly_heatmap %>% 
  layout(
    title = "Heapmap for Phosphrous Level over time by City"
  )
plotly_heatmap 
```

## Download PDF report

HTML report can be viewed [here](https://yuanxyyds.github.io/JSC370-Final-Project/Final_html.html)

Download PDF report [here](https://github.com/Yuanxyyds/JSC370-Final-Project/raw/main/Final_pdf.pdf)